Assessing paravascular transport in the parenchyma by PDE constrained optimization.

### 2018-10-22 tex/paper.tex
text

\begin{figure}
\includegraphics[width=0.95\textwidth]{Zoom-PatID-68.png} 
\caption{Shows the percentage intensity increase from baseline at different observation times in the slice (marked red in the left panel) used in the subsequent analysis.The colorbar was restricted to the range $(0,100)$}
\label{fig2} 
\end{figure}
Figure~\ref{fig1} shows the distribution of tracers during 24 hours, see also~\cite{ringstad2018brain} for further information on the imaging procedure.   
Our data (not all shown) consist of a total of 10 MRI observations, including a baseline MRI taken before tracer was injected. The observation points are distributed with 5 observations within 1-2 hours after injection, a single observation in the timeframes 2-4 hours, 6-9 hours, 24 hours and 48 hours. 
Figure~\ref{fig2} shows the region selected for our computations. 
The software Freesurfer \cite{Dale1999179, FischlLiuDale, spf2007, reuter:robreg10} was used to segment and align each of the observations, which made it possible to estimate voxelwise intensity increase.

The MRI data is obtained from a larger study, which invloves around 100 patients with similar MRI acquisitions. However, due to the clinical factor, some of the parameters involved are different. Thus a thorough analysis is needed to obtain a robust and accurate method.

@@

\beta  &= e\sp{- \sp{T_b}/_{T_1\sp{0}} } \\
\delta &= e\sp{- \sp{T_a}/_{T_1\sp{0}} } \\
\gamma &= e\sp{- \sp{T_w}/_{T_1\sp{0}} } \\
\rho   &= e\sp{- \sp{TR}/_{T_1\sp{0}}}  \\

@@

\begin{figure}
\centering
\includegraphics[width=0.70\textwidth]{T1function.png} 
\caption{Show the  function defined in \eqref{scaledmagnetization} with the white region indicating $T_1$ values for white matter,the grey region indicating $T_1$ values for grey matter,the blue region indicating $T_1$ values for CSF.  }
\label{figuredti} 
\end{figure}

\subsection{Diffusion tensor imaging}
In addition to the T1 and T2 weighted sequences described above we have also obtained diffusion tensor imaging (DTI) to assess the apparent diffusion 
coefficients on short time-scales. Images are shown in Figure~\ref{figuredti} where the larges diffusion coefficient (shown in 
red in the middle figure) is shown to be around $1.0e-3 mm^2/s$. We remark that we have not included possible anisotropy, shown in 
the right-most image in Figure~\ref{figuredti} and that these images shows the apparent diffusion coefficient for free water molecules (18 Da).      
The diffusivity of the Gadovist (600 Da \cite{MGadobutrol}) was estimated to be similar to the diffusion coefficient of Gd-DPTA (550 Da \cite{MGgDPTA}). This due to the fact that both molecules have similar mass, and based Stoke-Einstein equation should have similar diffusion coefficients. The free diffusion coefficient for Gd-DPTA was estimated in \cite{GdDPTA-DIFFUSION} to be $3.8e-4\mathrm{mm\sp{2}/s}$.
The fractional anisotropy is defined as 
\begin{equation}
FA\sp{2} =  \frac{3}{2} \frac{ (\lambda_1 - MD )\sp{2} +(\lambda_2 - MD )\sp{2} +(\lambda_3 - MD )\sp{2}}{\lambda\sp{2}_1 + \lambda\sp{2}_2  +\lambda\sp{2}_3 },
\end{equation}
with the mean diffusivity $MD$ defined as 
\begin{equation}
MD = \frac{\lambda_1 +\lambda_2 +\lambda_3 }{3}.
\end{equation}
In these equations $\lambda_i$ denotes the eigenvalues of the diffusion tensor.
\begin{figure}
\centering
\includegraphics[width=0.80\textwidth]{DTI-zoom.png} 
\caption{The left panel show the anatomical map. The middle panel show the apparent diffusion coefficients (ADC) obtained from DTI. The right panel shows the computed fractional anisotropy (FA) from the DTI.}
\label{figuredti} 
\end{figure}

@@

The mesh construction was patient-specific and used the MRI of a patient diagnosed with idiopathic normal pressure hydrocephalus (iNPH). The software Freesurfer was used in segmentation and creating the polyhedral surfaces of the white and grey matter. Then the use of T2 weighted MRI \cite{eidevalnes} was used to segment the CSF compartment surrounding the cerebral and the lateral ventricles. CGAL \cite{cgal:rty-m3-18b} was used to combine the polyhedral surfaces and to construed the mesh. The computational requirement for the resulting mesh was significant, therefore two submesh was also constructed, see Fig.\ref{Fig::Mesh}.
\begin{figure}
\centering
\includegraphics[scale=0.2]{mesh.png} 
\caption{A) Shows the mesh created from the baseline MR image with 3 domains. B) Shows the mesh created from the baseline MR image with 2 domains.  }
\label{Fig::Mesh}
\end{figure}
The 3 domain mesh in Fig.\ref{Fig::Mesh} A, consists of 244318 tetrahedral cells and 22057 vertices, while the 2 domain mesh Fig.\ref{Fig::Mesh} B consists of 335589 tetrahedral cells and 73002 vertices. The dimension of the mesh coordinates are $\mathrm{mm}$.

@@

The code used the FEniCS project (v.2017.2)  to solve the PDE, using backwards Euler time discretization and first order continuous Galerkin finite elements. The linear problem was solved using GMRES, and the matrix was stored in the solver. 
The module dolfin adjoint ~\cite{farrell2013automated} \kam{correct ref?}  was used to solve the PDE constrained optimization problem with the L-BFGS-B algorithm in the python module scipy \cite{LBFGSB1, LBFGSB2}.

The convergence criteria for the minimization was set so that 
\begin{equation}
||x ||_{L\sp{\infty}} \leq \epsilon
\end{equation} 
with $\epsilon=1.0e-1$ and $x$ denoting the projected gradient.

The observations $u_{obs}$ have a fixed time-points $t_i$, these time-points may not overlap with the time discretization in the forward problem $t_j$. Therefore a linear interpolation of the current solution $u(t_j)$ and previous solution $u(t_{j-1})$ was used approximate the solution at time $t_i$. The linear interpolation was given as  
\begin{equation}
\label{observation:interpolation}
u(t_i) \approx \frac{\Delta t}{dt} u(t_{j-1}) + \frac{dt - \Delta t }{dt} u(t_{j}), \quad t_i \in \lbrace t_{j-1}, t_j \rbrace
\end{equation}
with $\Delta t = t_j-t_i$ and $dt  = t_{j} - t_{j-1}$.

@@

The noise susceptibility was tested by adding a uniform distribution noise term to the observations. The noise term was constructed using the random library in numpy and adjusted so that noise range becomes $\lbrace -n_{amp} , n_{amp} \rbrace $, with $n_{amp}$ defined as noise amplitude.

\subsection{Acknowledgement}
This work was performed on the Abel Cluster, owned by the University of Oslo and Uninett\textbackslash Sigma2, and operated by the Department for Research Computing at USIT,the University of Oslo IT-department. \url{http://www.hpc.uio.no/}

@@

\subsection{Verification in terms of the manufactured solution}

Below we will discuss the parameter identification and its sensitivity with respect to regularization parameters, noise and time-resolution. In presentation of the results, the relative error is defined as 
\begin{equation}
 X\sp{rel} = \frac{X_{found} -X_{true} }{ X_{true} }
\end{equation}
with $X$ as an arbitrary scalar control parameter. For the boundary parameter $g$, the relative error norm is defined as 
\begin{equation}
|| g ||\sp{rel}_{L\sp{2}(\Omega_1)} = \frac{ \sum\limits_t\sp{k}|| g_{found}(t) -g_{true}(t)||_{L(\Omega_1)} }{  \sum\limits_t\sp{k}||g_{true}(t)||_{L(\Omega_1)} }
\end{equation}
In the case that the minimization algorithm fails to converge within 1000 iteration will be indicated by a hyphen in Table~\ref{TAB::timesteps}, Table~\ref{TAB::double} and Table~\ref{Tab::Noise0.3}

@@

\subsubsection{The number of timessteps}
Increasing the number of time-steps in the computations can potentially allow for a more accurate temporal reconstruction, but at the same time the number of observations per control decreases and the regularization may therefore play a larger role. Furthermore, the computational cost will rise with additional time-steps. Thus finding the optimal number of time-steps can be essential for larger problems. The computation varied the regularization parameters with 20 and 40 timesteps.

The results are given in Tab.\ref{TAB::timesteps}, and it can be seen for all $\alpha=1.0$ that the relative error becomes bigger with more time-steps. This behavior can also be seen for $\alpha=1.0e-2$ and $\beta < 1.0e-2$, which implies an interaction between the regularization parameters. 
Therefore $g$ was investigated by plotting boundary points through time for different values of $\alpha$ and $\beta$, see Fig \ref{boundarycontrol}. In Figure \ref{boundarycontrol} we can see that more time-steps causes  oscillations for $\alpha=1.0$, and each peak corresponds to a observation time. These oscillations are caused by $\alpha$, which tries minimize $g$ between observation times.

It can be seen that more time-steps causes an increase number of iterations, but for some values the number of iterations decreases. These values share that $\alpha $ is a magnitude 2 lower than $ \beta$,i.e $\alpha=1.0e-2$ and $\beta =1.0$. This indicates that $\alpha $ should be smaller than $\beta$ to achieve optimal convergence. However, the number of iterations in Table ~\ref{Tab::1} are smaller for all regularization parameters.

@@

The results are shown in Tab.\ref{TAB::double} and Tab.\ref{TAB::half}. In Tab.\ref{TAB::half}, the values $\alpha =1.0e-2$ and $\beta<1.0$ gives a surge in the relative error with more time-steps. This behavior is neither visible in Tab.\ref{TAB::timesteps} or Tab.\ref{TAB::double} given the same regularization parameters. It seems likely that this error is caused by oscillations, seen in Figure~\ref{boundarycontrol}.

The observations acts as stabilization for the functional, which gives a broader range of adequate regularization parameters.

@@

\subsubsection{The noise susceptibility}
The MRI data contains noise, hence an investigation to the noise susceptibility is needed. This was done by adding uniform noise in the range $\lbrace -n_{amp} , n_{amp} \rbrace $, with $n_{amp}$ defined as noise amplitude, to the manufactured observations. Then the optimization was solved by varying the regularization parameters $\alpha$ and $\beta$. Figure~\ref{12hourswithnoise} and Figure~\ref{24hourswithnoise} show the reconstruction of the manufactured solution with 
different noise levels with $\alpha=0.0001$ and $\beta=1.0$. Clearly, the reconstruction shown in the lower rows is quite robust with respect to the various levels
of noise shown in the upper rows.

In Tabel~\ref{Tab::Noise0.03}, we can see that a noise amplitude of 0.03, (10$\%$ of maximum initial condition) had negligible effect on the relative error. Then in Tabel~\ref{Tab::Noise0.3} the noise amplitude was increased to 0.3 (100$\%$ of maximum initial condition), which gave a visible effect. The noise addition caused the number of iteration to increase, and in combination with more timesteps caused the optimization not to convergence for $\beta < 1.0$. Furthermore, the relative boundary error $||g||\sp{rel}_{L\sp{2}(\Omega_1)}$ is significantly larger for $\alpha< 1.0e-2 $ and $\beta < 1.0e-2$.

@@

The median diffusion coefficient in the DTI was estimated to be $8.7e-4 \mathrm{mm\sp{2}/s}$ in white matter and $1.0e-3 \mathrm{mm\sp{2}/s}$ in grey matter. This corresponds to a tortuosity of $1.85$ and $1.73$, given that the self-diffusivity of water has been estimated to be around $3.0e-3\mathrm{mm\sp{2}/s}$ at $37\sp{o}C$. Fhe reference value $3.8e-4 \mathrm{mm\sp{2}/s}$ for Gd-DPTA gives an estimate for the apparent diffusion coefficient in the grey and white matter to respectively be $ 1.26e-4\mathrm{mm\sp{2}/s}$ and $1.10e-4 \mathrm{mm\sp{2}/s}$. This estimation assumes that the tortuosity is independent on molecular size.

The oscillation in Fig.\ref{boundarycontrol} showed a large discrepancy in the states between observations. Therefore the states with $k=48$ were examined after 36 hours, see Figure\ref{statecomparison}. It can be observed that the upper right row displays a discrepancy, which likely due to inadequate regularization parameters. Furthermore, we can see that for $\alpha=1.0e-2$ there is a clear discrepancy in the diffusion coefficients compared to other values $\alpha$. This prompt the exclusion of the corresponding values, which gives an average computed values $ 0.65 \mathrm{mm\sp{2}/h}$ and  $ 0.8 \mathrm{mm\sp{2}/h}$ in respectively grey and white matter. Scaled to $\mathrm{mm\sp{2}/s}$ gives the corresponding values $1.8e-4\mathrm{mm\sp{2}/s}$ and $2.22e-4 \mathrm{mm\sp{2}/s}$.

@@

In Fig.\ref{Fig::realdata}, it can be seen that there were regions with negligible amount of tracers and vice versa. The regions with above average amount of tracers corresponds with high diffusivity regions in Figure~\ref{figuredti}. This suggests that the computed diffusion coefficients corresponds to these values. The ADC value in the high diffusivity regions are closer to  $1.4e-3 \mathrm{mm\sp{2}/s}$, which corresponds to estimated diffusion coefficient of value $1.5e-3 \mathrm{mm\sp{2}/s}$. This gives a difference of $ 46 \%$ compared to the computed values.

@@

While a more comprehensive study involving more patients would be required in order to assess whether the clearance here happens faster than diffusion (\kam{egentlig er
vel dette det motsatte av clearance} a few remarks are in order. We find that the diffusion coefficients is 42-100\% larger than what was found by the DTI modality.  
However, we have not yet been able to assessment the self-diffusion of water as well as the diffusivity of Gadovist used in this study in phantom models. As such 
the constants in equation XXX must be taken with caution. Furthermore,    
the computational model assumes isotropic diffusivity, but the anisotropy in the white matter is well documented, as shown by the FA in Fig\ref{figuredti}. It can be seen in Figure~\ref{Fig::realdata} that the region with high FA have negligible amounts of tracer present. Therefore it would seem that the anisotropy do not have direct impact on the computations, since the computation can not evaluate a static environment. Therefore  observation of tracer in regions with anisotropy can be considered a requirement for adding anisotropy to the model.
The regions with high diffusivity in Figure~\ref{figuredti},  have significant amount of diffusivity as seen in Figure~\ref{Fig::realdata}

The computational model have 2 global controls for the diffusion coefficients, while it can be seen in  Fig.\ref{figuredti} that diffusion coefficients can be considered a spacial function. The implementation control parameters for each dofs would significantly increase the computational cost. The option using region specific control parameters seems to be a better option, since the diffusivity appears to be regional. It can also be taken into consideration to model the diffusion coefficients as a function in time, given the report of an increase clearance in rodents during sleeping \cite{xie2013sleep}. This would indicate a time-dependent diffusivity, since the clearance in the brain is driven by diffusion. The time-dependency would correspond to a change in tortuosity, since properties of the tracer molecule do not change.

>>>>

\begin{figure}
\includegraphics[width=0.95\textwidth]{Zoom-PatID-68.png} 
\caption{Shows the percentage intensity increase from baseline at different observation times in the slice (marked red in the left panel) used in the subsequent analysis.The color bar was restricted to the range $(0,100)$}
\label{fig2} 
\end{figure}
Figure~\ref{fig1} shows the distribution of tracers during 24 hours, see also~\cite{ringstad2018brain} for further information on the imaging procedure.   
Our data (not all shown) consist of a total of 10 MRI observations, including a baseline MRI taken before tracer was injected. The observation points are distributed with 5 observations within 1-2 hours after injection, a single observation in the timeframes 2-4 hours, 6-9 hours, 24 hours and 48 hours. 
Figure~\ref{fig2} shows the region selected for our computations. 
The software Freesurfer \cite{Dale1999179, FischlLiuDale, spf2007, reuter:robreg10} was used to segment and align each of the observations, which made it possible to estimate voxelwise intensity increase.

The MRI data was part of a larger study, which involves around 100 patients with the same MRI acquisitions. However, there are clinical differences in each patient, thus a thorough analysis is needed to obtain a robust and accurate method.

@@

\beta  &= e\sp{- \sp{T_b}/_{T_1\sp{c}} } \\
\delta &= e\sp{- \sp{T_a}/_{T_1\sp{c}} } \\
\gamma &= e\sp{- \sp{T_w}/_{T_1\sp{c}} } \\
\rho   &= e\sp{- \sp{TR}/_{T_1\sp{c}}}  \\

@@

\begin{figure}
\centering
\includegraphics[width=0.70\textwidth]{T1function.png} 
\caption{Show the  function defined in \eqref{scaledmagnetization} where the white region indicates $T_1$ values for white matter,the grey region indicates $T_1$ values for grey matter,the blue region indicates $T_1$ values for CSF.  }
\label{figuredti} 
\end{figure}

\subsection{Diffusion tensor imaging}
In addition to the T1 and T2 weighted sequences described above we have also obtained diffusion tensor imaging (DTI) to assess the apparent diffusion 
coefficients on short time-scales. Images are shown in Figure~\ref{figuredti} where the larges diffusion coefficient (shown in 
red in the middle figure) is shown to be around $1.0e-3 mm^2/s$. We remark that we have not included possible anisotropy, shown in 
the right-most image in Figure~\ref{figuredti} and that these images shows the apparent diffusion coefficient for free water molecules (18 Da).      
The diffusivity of the Gadovist (600 Da \cite{MGadobutrol}) was estimated to be similar to the diffusion coefficient of Gd-DPTA (550 Da \cite{MGgDPTA}). This due to the fact that both molecules have similar mass, and based on Stoke-Einstein equation should have similar diffusion coefficients. The free diffusion coefficient for Gd-DPTA was estimated in \citet{GdDPTA-DIFFUSION} to be $3.8e-4\mathrm{mm\sp{2}/s}$.
The fractional anisotropy is defined as 
\begin{equation}
FA\sp{2} =  \frac{3}{2} \frac{ (\lambda_1 - MD )\sp{2} +(\lambda_2 - MD )\sp{2} +(\lambda_3 - MD )\sp{2}}{\lambda\sp{2}_1 + \lambda\sp{2}_2  +\lambda\sp{2}_3 },
\end{equation}
with the mean diffusivity $MD$ defined as 
\begin{equation}
MD = \frac{\lambda_1 +\lambda_2 +\lambda_3 }{3}.
\end{equation}
In these equations $\lambda_i$ denotes the eigenvalues of the diffusion tensor.
\begin{figure}
\centering
\includegraphics[width=0.80\textwidth]{DTI-zoom.png} 
\caption{The left panel show the anatomical map. The middle panel show the apparent diffusion coefficients (ADC) obtained from DTI. The right panel shows the computed fractional anisotropy (FA) from the DTI.}
\label{figuredti} 
\end{figure}

@@

The mesh construction was patient-specific and used the MRI of a patient diagnosed with idiopathic normal pressure hydrocephalus (iNPH). The software Freesurfer was used in segmentation and creating the polyhedral surfaces of the white and grey matter. Then the use of T2 weighted MRI \cite{eidevalnes} was used to segment the CSF compartment surrounding the cerebral and the lateral ventricles. CGAL \cite{cgal:rty-m3-18b} was used to combine the polyhedral surfaces and to construct the mesh. The computational requirement for the resulting mesh was significant, therefore two submesh was also constructed, see Fig.\ref{Fig::Mesh}.
\begin{figure}
\centering
\includegraphics[scale=0.2]{mesh.png} 
\caption{A) Shows the mesh created from the baseline MR image with 3 domains. B) Shows the mesh created from the baseline MR image with 2 domains.  }
\label{Fig::Mesh}
\end{figure}
The 3 domain mesh in Fig.\ref{Fig::Mesh} A, consists of 244318 tetrahedral cells and 22057 vertices, while the 2 domain mesh Fig.\ref{Fig::Mesh} B consists of 335589 tetrahedral cells and 73002 vertices. The dimension of the mesh coordinates are $\mathrm{mm}$.

@@

The code used the FEniCS project (v.2017.2) to solve the PDE, using backwards Euler time discretization and first order continuous Galerkin finite elements. The linear problem was solved using GMRES, and the matrix was stored in the solver. 
The module dolfin adjoint ~\cite{farrell2013automated} \kam{correct ref?}  was used to solve the PDE constrained optimization problem with the L-BFGS-B algorithm in the python module scipy \cite{LBFGSB1, LBFGSB2}. The convergence criteria for the minimization was set so that 
\begin{equation}
||x ||_{L\sp{\infty}} \leq \epsilon
\end{equation} 
with $\epsilon=1.0e-1$ and $x$ denoting the projected gradient.

The observations $u_{obs}$ have a fixed time-points $t_i$, these time-points may not coincide  with the time discretization in the forward problem $t_j$. Therefore a linear interpolation of the current solution $u(t_j)$ and previous solution $u(t_{j-1})$ was used approximate the solution at time $t_i$. The linear interpolation was given as  
\begin{equation}
\label{observation:interpolation}
u(t_i) \approx \frac{\Delta t}{dt} u(t_{j-1}) + \frac{dt - \Delta t }{dt} u(t_{j}), \quad t_i \in \lbrace t_{j-1}, t_j \rbrace
\end{equation}
with $\Delta t = t_j-t_i$ and $dt  = t_{j} - t_{j-1}$.

@@

The noise susceptibility was tested by adding a uniform distribution noise term to the observations. The noise term was constructed using the random library in numpy and adjusted so that the noise range becomes $\lbrace -n_{amp} , n_{amp} \rbrace $, with $n_{amp}$ denoted as noise amplitude.

\subsection{Acknowledgements}
This work was performed on the Abel Cluster, owned by the University of Oslo and Uninett\textbackslash Sigma2, and operated by the Department for Research Computing at USIT,the University of Oslo IT-department. \url{http://www.hpc.uio.no/}

@@

\subsection{Verification in terms of the manufactured solution}
Below we will discuss the parameter identification and its sensitivity with respect to regularization parameters, noise and time-resolution. In presentation of the results, the relative error is defined as 
\begin{equation}
 X\sp{rel} = \frac{X_{found} -X_{true} }{ X_{true} }
\end{equation}
with $X$ as an arbitrary scalar control parameter. For the boundary parameter $g$, the relative error norm is defined as 
\begin{equation}
|| g ||\sp{rel}_{L\sp{2}(\Omega_1)} = \frac{ \sum\limits_t\sp{k}|| g_{found}(t) -g_{true}(t)||_{L(\Omega_1)} }{  \sum\limits_t\sp{k}||g_{true}(t)||_{L(\Omega_1)} }
\end{equation}
In the case that the minimization algorithm fails to converge within 1000 iteration will be indicated by a hyphen in Table~\ref{TAB::timesteps}, Table~\ref{TAB::double} and Table~\ref{Tab::Noise0.3}

@@

\subsubsection{The number of timessteps}
Increasing the number of time-steps in the computations can potentially allow for a more accurate temporal reconstruction, but at the same time the number of observations per control decreases and the regularization may therefore play a larger role. Furthermore, the computational cost will rise with additional time-steps. Thus finding the optimal number of time-steps can be essential for larger problems. The computation used a width range the regularization parameters, ans the number of timesteps were selected to be 20 and 40.

The results are given in Tab.\ref{TAB::timesteps}, and it can be seen for all $\alpha=1.0$ that the relative error becomes bigger with more time-steps. This behavior can also be seen for $\alpha=1.0e-2$ and $\beta < 1.0e-2$, which implies an interaction between the regularization parameters. Therefore $g$ was investigated by plotting boundary points through time for different values of $\alpha$ and $\beta$, see Fig \ref{boundarycontrol}. In Figure \ref{boundarycontrol} we can see that more time-steps causes oscillations for $\alpha=1.0$, and each peak corresponds to a observation time. These oscillations are caused by the value selection of $\alpha$, which tries minimize $g$ between observation times.

%It can be seen that more time-steps causes an increase number of iterations, but for some values the number of iterations decreases. These values share that $\alpha $ is a magnitude 2 lower than $ \beta$,i.e $\alpha=1.0e-2$ and $\beta =1.0$. This indicates that $\alpha $ should be smaller than $\beta$ to achieve optimal convergence. However, the number of iterations in Table ~\ref{Tab::1} are smaller for all regularization parameters.

@@

The results are shown in Tab.\ref{TAB::double} and Tab.\ref{TAB::half}. In Tab.\ref{TAB::half}, the values $\alpha =1.0e-2$ and $\beta<1.0$ gives a surge in the relative error with more time-steps. This behavior is neither visible in Tab.\ref{TAB::timesteps} or in  Tab.\ref{TAB::double} for the same regularization parameters. The ratio observations per control indicates that the increase in relative error is due to oscillations, similar to those seen in Figure~\ref{boundarycontrol}. Thus the observations acts as stabilization for the functional, which gives a broader range of adequate regularization parameters.

@@

\subsubsection{The noise susceptibility}
The MRI data contains noise, hence an investigation to the noise susceptibility is needed. This was done by adding uniform noise in the range $\lbrace -n_{amp} , n_{amp} \rbrace $, with $n_{amp}$ defined as noise amplitude, to the manufactured observations. Then the optimization was solved by varying the regularization parameters $\alpha$ and $\beta$. Figure~\ref{12hourswithnoise} and Figure~\ref{24hourswithnoise} shows the reconstruction of the manufactured solution at
different noise levels with $\alpha=0.0001$ and $\beta=1.0$. Clearly, the reconstruction shown in the lower rows is quite robust with respect to the various levels
of noise shown in the upper rows.

In Tabel~\ref{Tab::Noise0.03}, we can see that a noise amplitude of 0.03, (10$\%$ of maximum initial condition) had negligible effect on the relative error. Then in Tabel~\ref{Tab::Noise0.3} the noise amplitude was increased to 0.3 (100$\%$ of maximum initial condition), which gave a visible effect. The noise caused the number of iteration to increase, and with more timesteps caused the optimization not to convergence for $\beta < 1.0$. Furthermore, the relative boundary error $||g||\sp{rel}_{L\sp{2}(\Omega_1)}$ is significantly larger for $\alpha< 1.0e-2 $ and $\beta < 1.0e-2$.

@@

The median diffusion coefficient in the DTI was estimated to be $8.7e-4 \mathrm{mm\sp{2}/s}$ in white matter and $1.0e-3 \mathrm{mm\sp{2}/s}$ in grey matter. This corresponds to a tortuosity of $1.85$ and $1.73$, given that the self-diffusivity of water has been estimated to be around $3.0e-3\mathrm{mm\sp{2}/s}$ at $37\sp{o}C$. Then the reference value $3.8e-4 \mathrm{mm\sp{2}/s}$ for Gd-DPTA gives an estimate for the apparent diffusion coefficient in the grey and white matter to  be $ 1.26e-4\mathrm{mm\sp{2}/s}$ and $1.10e-4 \mathrm{mm\sp{2}/s}$ respectively. This estimation assumes that the tortuosity is independent on molecular size.

The oscillation in Fig.\ref{boundarycontrol} showed a large discrepancy in the states between observations. Therefore the states with $k=48$ were examined after 36 hours, see Figure\ref{statecomparison}. It can be observed that the upper right row displays a discrepancy, which likely due to inadequate regularization parameters. Furthermore, we can see that for $\alpha=1.0e-2$ there is a clear discrepancy in the diffusion coefficients compared to other values $\alpha$. This prompt the exclusion of these values, which gives an average computed values $ 0.65 \mathrm{mm\sp{2}/h}$ and  $ 0.8 \mathrm{mm\sp{2}/h}$ in grey and white matter. Scaled to $\mathrm{mm\sp{2}/s}$ gives the corresponding values $1.8e-4\mathrm{mm\sp{2}/s}$ and $2.22e-4 \mathrm{mm\sp{2}/s}$.

@@

In Fig.\ref{Fig::realdata}, it can be seen that there were regions with negligible amount of tracers and vice versa. The regions with above average amount of tracers corresponds with high diffusivity regions in Figure~\ref{figuredti}. This suggests that the computed diffusion coefficients corresponds to these values. The ADC value in the high diffusivity regions are approximately  $1.4e-3 \mathrm{mm\sp{2}/s}$, which corresponds to estimated diffusion coefficient of value $1.5e-3 \mathrm{mm\sp{2}/s}$. This gives a difference of $ 46 \%$ compared to the computed values.

@@

While a more comprehensive study involving more patients would be required in order to assess whether the clearance here happens faster than diffusion (\kam{egentlig er
vel dette det motsatte av clearance} a few remarks are in order. We find that the diffusion coefficients is 42-100\% larger than what was found by the DTI modality.  
However, we have not yet been able to assessment the self-diffusion of water as well as the diffusivity of Gadovist used in this study in phantom models. As such 
the constants in equation XXX must be taken with caution. Furthermore,    
the computational model assumes isotropic diffusivity, but the anisotropy in the white matter is well documented, as shown by the FA in Fig\ref{figuredti}. It can be seen in Figure~\ref{Fig::realdata} that the region with high FA have negligible amounts of tracer present. Therefore it would seem that the anisotropy do not have direct impact on the computations, since the computation can not evaluate a static environment. Therefore  observation of tracer in regions with anisotropy can be considered a requirement for adding anisotropy to the model.

The computational model uses 2 global controls for the diffusion coefficients, while it can be seen in  Fig.\ref{figuredti} that diffusion coefficients can be considered a spacial function. The implementation control parameters for each degree of freedom, would significantly increase the computational cost. The option using region specific control parameters seems to be a better option, since the diffusivity appears to be regional. It can also be taken into consideration to model the diffusion coefficients as a function in time, given the report of an increase clearance in rodents during sleeping \cite{xie2013sleep}. This would indicate a time-dependent diffusivity, since the clearance in the brain is driven by diffusion. The time-dependency would correspond to a change in tortuosity, since properties of the tracer molecule do not change.
